---
title: "data manipulation"
output: github_document
---

# load in the FAS litters data
```{r setup, we want readr}
library(tidyverse)
library(readxl)
litters_df = read_csv("./data/FAS_litters.csv", na = c("NA", ".", ""))
litters_df = janitor::clean_names(litters_df)
```

# select
choose colums of group, litter_number, gd0_weight, pups_born_alive in dataset litters_df.
```{r}
select(litters_df, group, litter_number, gd0_weight, pups_born_alive)
```
from cloum ... to...
```{r}
select(litters_df, group:gd_of_birth)
```
get gid of ...
```{r}
select(litters_df, -pups_survive)
```
rename colum, and keep others
```{r}
select(litters_df, GROUP = group, LiTtEr_NuMbEr = litter_number)
```
keep colums start with ...
```{r}
select(litters_df, starts_with("gd"))
```
move specific columns to the front, keep everything else remain
```{r}
select(litters_df, litter_number, pups_survive, everything())
```
move column to front
```{r}
relocate(litters_df, litter_number, pups_survive)
```

# filter
!() filters that () is not ture
```{r}
filter(litters_df, gd_of_birth <= 22)
filter(litters_df, !(pups_survive == 4), group %in% c("Con7", "Con8"))
```


# mutate
### creat new variables

```{r}
mutate(litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```

# arrange
### put things in order

order group from small to large, then pups_born_alive
```{r}
arrange(litters_df, group, pups_born_alive)
```

# pipe %>% or |>

not use pipe
```{r}
litters_df_raw = read_csv("./data/FAS_litters.csv", na = c("NA", ".", ""))
litters_df_clean_names = janitor::clean_names(litters_df_raw)
litters_df_selected_cols = select(litters_df_clean_names, -pups_survive)

litters_df_with_vars = 
  mutate(
    litters_df_selected_cols, 
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group))
litters_df_with_vars_without_missing = 
  drop_na(litters_df_with_vars, wt_gain)
litters_df_with_vars_without_missing


litters_df_clean = 
  drop_na(
    mutate(
      select(
        janitor::clean_names(
          read_csv("./data/FAS_litters.csv", na = c("NA", ".", ""))
          ), 
      -pups_survive
      ),
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)
    ),
  wt_gain
  )
```

Use pipe to achieve above result:
```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv", na = c("NA", ".", "")) |> 
  janitor::clean_names() %>%
  select(-pups_survive) |> 
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>%
  drop_na(wt_gain)
```




