---
title: "data import"
output: github_document
---


```{r setup, we want readr}
library(tidyverse)
library(readxl)
```

# read in some data

Use reletive path, so people download from github can also access the dataset.
./ means start from this github "data_wargling" folder.
Since we don't use other function in janitor, so that we don't need to write library(janitor).
```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

# take a look at dataset

whole dataset:
```{r}
litters_df
```
first 6 rows:
```{r}
head(litters_df)
```
last few rows:
```{r}
tail(litters_df)
```
few statistics data:
```{r}
skimr::skim(litters_df)
```

# options to read_csv
skip=10 means skipping 10 rows. col.names=FALSE is to make up colum names as x1 x2 x3... rather than the origenal colum name
```{r}
?read_csv
```

# read excel file
```{r}
mlb_df = read_excel("./data/mlb11.xlsx", range = "A1:F7")
mlb_df
```

# read SAS file
```{r}
library(haven)
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
head(pulse_df, 5)
```
#Comparison whith Base R
###don't use 'read.csv'

```{r}
pups_base = read.csv("./data/FAS_pups.csv")
pups_readr = read_csv("./data/FAS_pups.csv")

View(pups_base)
View(pups_readr)

pups_base
pups_readr

pups_base$S
pups_readr$S
```


# export data
```{r}
write_csv(mlb_df, "./data/mlb_subtable.csv")
```