---
title: "data tidy"
output: github_document
---

## load in the PULSE data in sas format
```{r setup, we want readr}
library(tidyverse)
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") |>
  janitor::clean_names()
```

## make the wide table a long table
Put column name "bdi_score_bl:bdi_score_12m" into "visit", value in these colum into "bdi". And delete all "bdi_score_" in every "visit" value.
```{r}
pulse_data_tidy = pulse_data %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi"
    )

pulse_data_tidy
```

Combine previous steps together, and add mutate:
```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") |>
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi"
    ) %>%
  relocate(id, visit) %>%
  mutate(visit = recode(visit, "bl" = "00m"))
```

## pivot_wider

make up new dataset "analysis_result"

pivot_wider is the inverse of pivot_longer
```{r}
analysis_result = 
  tibble(
    group = c("treatment", "treatment", "placebo", "placebo"),
    time = c("pre", "post", "pre", "post"),
    mean = c(4, 8, 3.5, 4)
  )

analysis_result %>% 
  pivot_wider(
    names_from = "time", 
  values_from = "mean"
  )
```

## Binding rows

using LotR data.
1. import each table
```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_words.xlsx", range = "B3:D6") %>%
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data/LotR_words.xlsx", range = "F3:H6") %>%
  mutate(movie = "fellowship_ring")

return_king = 
  readxl::read_excel("./data/LotR_words.xlsx", range = "J3:L6") %>%
  mutate(movie = "fellowship_ring")
```
2. bind 3 tables together
```{r}
lotr_tidy = 
  bind_rows(fellowship_ring, two_towers, return_king) |>
  janitor::clean_names() |>
  pivot_longer(
    female:male,
    names_to = "gender", 
    values_to = "words") |>
  mutate(race = str_to_lower(race)) |> 
  select(movie, everything()) 

```

## Joining datasets
import FAS datasets
separate(group, into = c("dose", "day_of_tx"), sep = 3) means in group colum, first 3 character go to "dose", other go to "day_of_tx".
```{r}
pup_df = 
  read_csv("./data/FAS_pups.csv", na = c("NA", "", ".")) |>
  janitor::clean_names() |>
  mutate(sex = recode(sex, "1" = "male", "2" = "female")) 

litter_df = 
  read_csv("./data/FAS_litters.csv", na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  relocate(litter_number) |>
  separate(group, into = c("dose", "day_of_tx"), sep = 3) 
  
```
then join them
```{r}
fas_df = 
  left_join(pup_df, litter_df, by = "litter_number") %>%
  arrange(litter_number) %>%
  relocate(litter_number, dose, day_of_tx)
```

